---
title: "NFL Data Part 1: QBs"
output: html_notebook
---

```{r Loading Packaages, message=FALSE, warning=FALSE}
packages <- c("tidyverse", "lubridate","forcats","gt","devtools","ggimage",
              "knitr","Hmisc","psych","summarytools","DT","nflscrapR","nflfastR",
              "glue","tidyquant","rvest")

if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())))  
}else ("all packages installed")
for (i in packages) {
  library(i, character.only = TRUE)
}
rm(packages,i)

# devtools::install_github(repo = "maksimhorowitz/nflscrapR")
# devtools::install_github("mrcaseb/nflfastR")

knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)

mode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
```

Loading in Play by Play data for 2015-2019

```{r Querying the Data, message=FALSE, warning=FALSE}
seasons <- 2010:2019
pbp <- map_df(seasons, function(x) {
  readRDS(
    url(
      paste0("https://raw.githubusercontent.com/guga31bb/nflfastR-data/master/data/play_by_play_",x,".rds")
    )
  )
})

pbp
```

Gathering Team information

```{r Team Data}

team_info = teams_colors_logos
team_info

```

Play-by_Play data for QBs with relevant columns

```{r Wrangling QB PBP Data, message=FALSE, warning=FALSE}
pbp_qb = pbp %>% 
  filter(pass_attempt == 1, aborted_play != 1,play_deleted == 0) %>% 
  select(posteam,season,week,
    # Player Info
    passer_name = passer_player_name,passer_id = passer_player_id, receiver_name = receiver,receiver_id = receiver_player_id,
    # Yardage Info
    yards_gained, air_yards, yac = yards_after_catch, 
    # Completion Info
    first_dn = first_down_pass, complete_pass,incomplete_pass, interception, pass_touchdown,
    # Non-Completion Info
    qb_hit,sack,fumble,fumbled_1_player_id,
    #Pay Type Infor
    shotgun,no_huddle) %>% 
  arrange(-season)

pbp_qb
```

Summary Information for the QBs. Grouping by season and filtering out the Playoffs.

```{r Summarise the QB data ,message=FALSE, warning=FALSE}
qbs = pbp_qb %>% 
  filter(week <= 17) %>% 
    mutate(fumble = ifelse(fumbled_1_player_id == passer_id ,1,          0),
    sack_yards    = ifelse(sack   == 1,                     yards_gained,NA),
    yards_gained  = ifelse(sack   == 1,                     0,           yards_gained)) %>% 
  group_by(passer_name, passer_id,season) %>% 
  summarise_if(is.numeric, list(sum), na.rm = T) %>% 
  mutate(att = complete_pass+incomplete_pass + interception) %>% 
  select(passer_name,season, cmp = complete_pass,att,yds = yards_gained,air_yds = air_yards,yac,
         TD = pass_touchdown,INT = interception,first_dn,everything(),sack_yds = sack_yards) %>% 
    ungroup() %>% 
  select(-passer_id,-week,-incomplete_pass) %>% 
  arrange(-att) %>% 
  filter(att > 100)
  
qbs
```

Normalizing the QB data based on the number of attempts. Sack yardage data gets normalized to the number of sacks.

```{r Normalizing the data}
qbs_perAtt = qbs %>%
  mutate( across(-c(passer_name,season,att,sack,sack_yds), 
                 .fns = list(perAtt = ~ ./att)    ),
          yds_perSack = sack_yds/sack,
          sack_perAtt = sack/att
          ) %>% 
  select(passer_name,season,att,contains('per')) %>% 
  mutate( across(-c(passer_name,att,season),
                 .fns = list(~round(.,3)),
                 .names = "{col}"             )
  ) %>% 
  arrange(-air_yds_perAtt)

qbs_perAtt_now = qbs_perAtt%>% 
  filter(season == max(season))


datatable(qbs_perAtt_now, filter = 'top', options = list(pageLength = 25, autoWidth = F,scrollX=TRUE))
```

Repeating the same basic steps for recievers (regardless of position).

```{r Cleaning and Wrangling Receiver Data, message=FALSE, warning=FALSE}
pbp_receiver = pbp %>% 
  mutate(fumble = ifelse(fumbled_1_player_id == receiver_id ,1,          0)) %>% 
  filter(pass_attempt == 1, aborted_play != 1,play_deleted == 0, !is.na(receiver_id)) %>% 
  select(posteam,season,week,down,
    # Player Info
    receiver_name = receiver,receiver_id = receiver_player_id,
    # Yardage Info
    yds = yards_gained, air_yards, yac = yards_after_catch, 
    # Completion Info
    first_dn = first_down_pass,TD = pass_touchdown,complete_pass,
    # Non-Completion Info
    fumble) %>% 
  arrange(-season)

receiver = pbp_receiver %>% 
  filter(week <= 17) %>% 
  group_by(receiver_name,receiver_id,season) %>% 
  summarise(
    team = last(posteam),
    targets = n(),
    rec = sum(complete_pass, na.rm = T),
    yds = sum(yds, na.rm = T),
    air_yds = sum(air_yards, na.rm = T),
    yac = sum(yac, na.rm = T),
    first_dn = sum(first_dn, na.rm = T),
    TD = sum(TD, na.rm = T),
    fumble = sum(fumble, na.rm = T)
  ) %>% 
  ungroup() %>% 
  select(-receiver_id) %>% 
  arrange(-yds)

receiver_now = receiver %>% 
  filter(season == max(season))

receiver_now
```

```{r message=FALSE, warning=FALSE}
receiver_perCatch = receiver %>%
  mutate( across(-c(receiver_name,season,team,targets,rec), 
                 .fns = list(perCatch = ~ round(./rec,3))   ),
          'catch_%' = round(rec/targets,3)
          ) %>% 
  select(receiver_name,season,team,rec,'catch_%',contains('per')) %>% 
  arrange(-rec)


receiver_perCatch_now = receiver_perCatch %>% 
  filter(season == max(season))


datatable(receiver_perCatch_now, filter = 'top', options = list(pageLength = 25, autoWidth = F,scrollX=TRUE))

```

```{r}



```
